##############################################################################################
@title 
geoMorphThalwegMetrics

@author
Rachel Krauss \email{rkrauss@battelleecology.org}

@description 
This script calculates the following thalweg metrics 
1. Total thalweg length (m)
2. Thalweg length by habitat ID
3. Percent length by habitat type
4. Thalweg length between S1 and S2  
5. Habitat units between S1 and S2 
6. Slope for the thalweg and between S1 and S2
7. Percentage of dry reach using thalweg lengths (if applicable)
8. Area of the habitat units
9. Number of geomorphology features

The code then creates shapefiles with standardized formatting

@parmam
All the inputs are arcGIS shapefiles created from the raw total station survey data and trimble GPS files. See "set your variables" for descriptions of each input

@return
1.Six-seven csvs depending on the inputs that contain the thalweg metrics listed in the description
2.Four- six shapefiles depending on the inputs

@changelog
R. Krauss (2017-12-11)
original creation
##############################################################################################
```{r packages}
#load packages
require(dplyr)
require(plyr)
require(sp)
require(sf)
require (rgdal)
require(rgeos)
require(raster)
```

```{r set your variables}
filePath <- "N:/Science/AQU/Geomorphology_Survey_Data/D08/MAYF_20171011/processedData" #location where the shapefiles are stored 
outPath <- "N:/Science/AQU/Geomorphology_Survey_Data/D08/MAYF_20171011/processedData/outputsR"  #output Location
site <- "MAYF" #4 letter site code
domain <- "D08" #DXX,for example D01, D11
surveyDate <- "20171103" # YYYYMMDD see N:\Science\AQU\Geomorphology_Survey_Data\Geomorph_Map_Progress.xlsx


#files you always need to include

thalwegShapefileName <- "thalweg"  #thalweg polyline shapefile
habitatShapefileName <- "habitat"  #habitat polygon shapefile
surveyShapefileName <- "processedPts"  #survey point shapefile that includes s1 & s2, staff gauge, and transect points
sensorBoxShapefileName <- "sensorBox" #polygon shapefile of area in between the sensors

#files that may or may not exisit at the site.  Run this lines anyways, even if the files do not exist
geoFeaturesShapefileName <- "geomorphicFeatures" #polygon shapefile that includes of MCBs, ISLs, downed logs, beaver dams

#Write the name of the thalweg point furthest upsteam (some sites will be downstream, this is ok, the code takes the abs slope)
# see N:\Science\AQU\Geomorphology_Survey_Data\lastThalwagShots.csv
upStreamTHL <- "THL385"
```

```{r load the shapefiles}
#to use readOGR you need to separate the wd and the specific shapefile
thalweg <- readOGR(filePath,thalwegShapefileName)
rawHabitat <- readOGR(filePath,habitatShapefileName)
surveyPts <- readOGR(filePath,surveyShapefileName)
sensorBox <-readOGR(filePath,sensorBoxShapefileName)
if (file.exists(paste(paste(filePath,geoFeaturesShapefileName, sep="/"),".shp",sep=""))){
  geoFeatures <- readOGR(filePath,geoFeaturesShapefileName)
}

#plot the shapefiles to make sure you loaded the correct ones
plot(thalweg)
plot(rawHabitat, add=TRUE)
plot(sensorBox, add =TRUE)
plot(surveyPts, add= TRUE)
if (file.exists(paste(paste(filePath,geoFeaturesShapefileName, sep="/"),".shp",sep=""))){
  plot(geoFeatures, add= TRUE)
}
```

```{r remove the MCB and ISL from the habitat area}
#erases the mcb,island, beaver dams, etc area from the habitat id polygon if they exist 
if (file.exists(paste(paste(filePath,geoFeaturesShapefileName, sep="/"),".shp",sep=""))){
  #remove the areas from the habitat Id layer that include geomorpth features
  habitat2 <- raster::erase(rawHabitat,geoFeatures)

  #check to make sure the features are removed- there should now be holes in the habitat layer
  plot(habitat2)
  
  #calculate area
  habitat2@data$area_m2 <- gArea(habitat2, byid=TRUE)
  
} else { #if no geomorphology features exist
  #rename habitat to habitat 2 for the rest of the code
  habitat2 <- rawHabitat
  
  #calculate area
  habitat2@data$area_m2 <- gArea(habitat2, byid=TRUE)
}

#check to make sure area was added
habitat2@data
```

```{r braided channels} 
#this chunk compares the length of each channel where braided channels occur and keeps the longest channel to produce the long profile line

#add length (in m) to each line segment in the thalweg line shapefile
thalweg@data$length <- gLength(thalweg, byid=TRUE)

#compare the lengths for the braided channels and choose one to keep
#filter out all the lines that have a digit in the feature field (mcb1, etc)
geoFeaturesOnly <- thalweg[grepl("\\d", thalweg@data$featureID),] 

#check the data- only the channels should plot
geoFeaturesOnly@data
plot(geoFeaturesOnly)

#for each geo feature the line below compares the 2 lengths and then keeps the max
geoFeatMax <- geoFeaturesOnly@data %>% group_by(featureID) %>% filter(length == max(length))
geoFeatMax <- data.frame(geoFeatMax)

#you cannot use dplyr on spatial attribute tables so I created a list with the mcb rows I wanted to keep and then filtered the attribute table by that list
temp <- paste(geoFeatMax$featureID,geoFeatMax$length)
geoFeaturesOnly@data$temp <- paste(geoFeaturesOnly@data$featureID,geoFeaturesOnly@data$length)
geoFeaturesRevised<- geoFeaturesOnly[(geoFeaturesOnly@data$temp %in% temp),]

#remove the temp column 
geoFeaturesRevised <- geoFeaturesRevised[,-5]

#check the data, should be just one line for each mcb
geoFeaturesRevised@data
plot(geoFeaturesRevised)

#remove all mcb lines from the original dataframe and add back in the ones we want to keep
thalweg2 <- thalweg[!grepl("\\d", thalweg@data$featureID),]
thalwegUpdatedGeoFeatures <- rbind(thalweg2,geoFeaturesRevised)

#update any site specific changes
#pringle had a thalweg terminate into a bank, remove that section
if (site == "PRIN"){
  thalwegUpdatedGeoFeatures <- thalwegUpdatedGeoFeatures[!grepl("terminate", thalwegUpdatedGeoFeatures@data$comment),]
}

#now the thalweg line should contain no braided channels (ie, one line across the reach)
thalwegUpdatedGeoFeatures@data
plot(thalwegUpdatedGeoFeatures)
```

```{r intersect thalweg, SIS2 by habitat type}
#intersect the thalweg line by the habitat polygon.  
#for each habitat id (riffle 1, riffle 2, etc) this code breaks up the thalweg line to match those polygons
habitatOverlay <- raster::intersect(thalwegUpdatedGeoFeatures, rawHabitat)

#add length (in m) 
habitatOverlay@data$LengthHab <- gLength(habitatOverlay, byid=TRUE)

#create a line shapefile between S1 & S2 
S1S2Overlay <- raster::intersect(thalwegUpdatedGeoFeatures, sensorBox)
plot(S1S2Overlay)
S1S2Overlay@data$Length <- gLength(S1S2Overlay, byid=TRUE)

#determine what habitat types S1 and S2 are in 
#pull out S1 and S2 from the point shapefile-  watch out to make sure you only pull 2 points
S1S2 <- surveyPts[grepl("^S1$|^S2$",surveyPts@data$name),]
plot(S1S2)
plot(rawHabitat, add=T)

#overlay the habitat polygon
S1S2Habitat <- raster::intersect(S1S2, rawHabitat)
```

```{r create the thalweg long profile}
#this chunk pulls out the thalweg points that fall along the updated thalweg line

#first create a small buffer to transform the thalwg line into a polygon
thalBuff<- gBuffer(thalwegUpdatedGeoFeatures, width= 2)

#pull out all thalweg and max pool depth points from the surveys
thalPoints <- surveyPts[grepl("THL|MPD",surveyPts@data$name),]

#overlay the survey points with the modified thalweg
longProfile <- raster::intersect(thalPoints, thalBuff)

#check data
plot(longProfile)
longProfile@data

#convert to data.frame to get ready for export
longProfileDF <- data.frame(longProfile)

#clean up the dataframe columns
longProfileDF <- longProfileDF[,-c(8:11)]

#change transect names back to thalweg
longProfileDF$mapCode <- ifelse(grepl("Transect",longProfileDF$mapCode),"THL",as.character(longProfileDF$mapCode))
```


```{r calculations}
#first part calcuates the entire length of thalweg by habitat types
thalwegHabitat<- data.frame(habitatOverlay)
thalwegHabitat %>% mutate_if(is.factor, as.character) -> thalwegHabitat

#clean up the dataframe by dropping fields we no longer need
thalwegHabitat <- thalwegHabitat[,-c(1:4,6)]

#add all the lengths up for each habitat ID.  
#This line says for the thalweg habitat dataframe, groupy by Habitat ID (riffle1, riffle 2, etc)
#and then sum the length of all the lines segements that happen in that area.
habitatID <- ddply(thalwegHabitat , .(habitatID), summarise, lengthM= sum(LengthHab))

# add back the area information (use habitat2 which has removed area from geomorph features)
habitatID$area_m2 <-0
habitat2DF <- data.frame(habitat2)

for (a in 1:nrow(habitatID)){
  for (z in 1:nrow(habitat2DF)){
  if(habitatID$habitatID [a] == habitat2DF$habitatID [z]){
    habitatID$area_m2[a] <- habitat2DF$area_m2[z]
    }
  }
}

#Compute % length by habitat ID and then by habitat type
totalLength <- sum(habitatID$lengthM)
totalArea <- sum(habitatID$area_m2)
habitatID$percentLength <-habitatID$lengthM/totalLength*100

#add totals to the bottom of the dataframe
myNumCols <- which(unlist(lapply(habitatID, is.numeric)))
habitatID[(nrow(habitatID) + 1), myNumCols] <- colSums(habitatID[, myNumCols], na.rm=TRUE)

#create habitat type summary table
habitatID$habitat <- gsub("\\d",'\\1',habitatID$habitatID)
habitatLength<- ddply(habitatID, .(habitat), summarise, lengthM= sum(lengthM))
habitatArea<- ddply(habitatID, .(habitat), summarise, areaM2= sum(area_m2))

#combine information into one dataframe
habitat <- merge(habitatLength,habitatArea, by='habitat')

#calculate percents for each habitat table
habitat$percentLength <-habitat$lengthM/totalLength*100

#clean up the processed points
#transform to a dataframe
surveyPtsDF <- data.frame(surveyPts)

#clean up the processed survey points
surveyPtsDF <- surveyPtsDF[,-c(8:11)]

#calculate slope of the thalweg
#downstream elevation = THL
thalwegLength <- sum(habitatID$dryLengthM, na.rm=TRUE)
dsElevation <- surveyPtsDF$elevation[which(surveyPtsDF$name == "THL")]
usElevation <- surveyPtsDF$elevation[which(surveyPtsDF$name == upStreamTHL)]

habitat$slope <-ifelse(is.na(habitat$habitat),round(abs(usElevation-dsElevation)/totalLength,4),"")

#calculate the length between S1 and S2, slope, and the assoicated habitat types
S1S2thalweg <- data.frame(S1S2Overlay)
S1S2DF<- data.frame(S1S2)
S1S2Metrics<- data.frame(length=(sum(S1S2thalweg$Length)), elevationDiff = S1S2DF$elevation[which(S1S2DF$name == "S1")]- S1S2DF$elevation[which(S1S2$name == "S2")])
S1S2Metrics$slope <- S1S2Metrics$elevationDiff/S1S2Metrics$length

S1S2HabitatDF <- data.frame(S1S2Habitat)
S1S2habitat <- data.frame(sensor =S1S2HabitatDF$name,  easting=S1S2HabitatDF$easting, northing =S1S2HabitatDF$northing,latitude= S1S2HabitatDF$latitude, longitude=S1S2HabitatDF$longitude, elevation= S1S2HabitatDF$elevation,habitatID =S1S2HabitatDF$habitatID)
#you will get an error if S1/S2 have different names

#geomorph feature counts
if (file.exists(paste(paste(filePath,geoFeaturesShapefileName, sep="/"),".shp",sep=""))){
  geoFeaturesDf <- data.frame(geoFeatures)
  #pull out first 3 letters 
  geoFeaturesDFSUM<- data.frame(feature = as.character(substr(geoFeaturesDf$featureID,1,3)))
}

#pull out any waterfalls, culverts, tribs, from the survey points
surveyFeats <- data.frame(surveyPts)
surveyFeats2 <- surveyFeats[grepl("WFT|CUL|TRB|TRI|MAR|WEIR_L",surveyFeats$name),] 
surveyFeats2$name[which(surveyFeats2$mapCode=="Waterfall")]<- "WF" #waterfalls have 2 points, top and bottom.
surveyFeats2$name[which(surveyFeats2$mapCode=="Weir")]<- "WEIR" #weir has 2 points, left and right.

if (nrow(surveyFeats2) > 0){
  #pull out first 3 letters 
  surveyFeatsSUM<- data.frame(feature = as.character(substr(surveyFeats2$name,1,3)))
  surveyFeatsSUM$feature[which(surveyFeatsSUM$feature=="WEI")]<- "WEIR" #add the "r" back in
}

if (nrow(geoFeaturesDf) > 0 & nrow(surveyFeats2) > 0){
  featSum <- rbind(geoFeaturesDFSUM,surveyFeatsSUM)
  featSummary <- plyr::count(featSum, "feature")
} else if ((nrow(geoFeaturesDf) > 0) & (nrow(surveyFeats2) == 0)){
  featSummary <- plyr::count(geoFeaturesDFSUM, "feature")
}else if (nrow(surveyFeats2) > 0 & nrow(geoFeaturesDf) == 0){
  featSummary <- plyr::count(surveyFeatsSUM, "feature")
} else {
  print("No geomorph features or feature points to count")
}
```

```{r dry sections}
#calculates the # of the thalweg that is dry
dryDF <- data.frame(habitatOverlay)

if ("Y" %in% dryDF$dryComment){
  #cleanup DF
  dryDFHabitat <- dryDF[,-c(1,3,4,6)]
  #add all the lengths up for each habitat ID.  
  #This line says for the thalweg habitat dataframe, group by Habitat ID (riffle1, riffle 2, etc)and dry comment
  #and then sum the length of all the lines segments that happen in that area.
  dryDFHabitat <- ddply(dryDFHabitat , .(habitatID, dryComment), summarise, dryLengthM= sum(LengthHab))
  #calculate % dry total
  dryDFHabitat$dryLengthM <- round(dryDFHabitat$dryLengthM,2)
  #drop any that round to 0
  dryDFHabitat <- dryDFHabitat[!(dryDFHabitat$dryComment == "Y" & dryDFHabitat$dryLengthM ==0.00),]
  thalwegLength <- sum(dryDFHabitat$dryLengthM, na.rm=TRUE)
  thalwegLengthDry <- sum(dryDFHabitat$dryLengthM[which(dryDFHabitat$dryComment== "Y")])
 
  #add in the percentdry per habitat ID
   dryLength <- filter(dryDFHabitat, dryComment == "Y")
   habitatID<- left_join(habitatID,dryLength, by ="habitatID") 
   # calute percent dry
   habitatID$percentDry <- round(habitatID$dryLengthM/habitatID$lengthM*100,2)
   #change any values over 100 back to 100 (happens due to rounding error)
   habitatID$percentDry <-ifelse(habitatID$percentDry > 100, 100.00, habitatID$percentDry)
  #Add the totaldry length and the total precent dry
   habitatID$percentDry <-ifelse(is.na(habitatID$habitatID),round(thalwegLengthDry/thalwegLength*100,2),habitatID$percentDry)
   habitatID$dryLengthM<-ifelse(is.na(habitatID$habitatID),thalwegLengthDry,habitatID$dryLengthM)
   
   
  #clean up and reorder the dataframe
   habitatID<- data.frame(habitatID= habitatID$habitatID, lengthM=habitatID$lengthM,
                         area_m2= habitatID$area_m2, percentLength= habitatID$percentLength,
                         dryLengthM= habitatID$dryLengthM,
                         percentDry=habitatID$percentDry, 
                         habitat= habitatID$habitat, stringsAsFactors = FALSE)
   #change NA to blank
   habitatID[is.na(habitatID)] <- ""
}
```

```{r write outputs}
#round all the output dataframes
round_df <- function(x, digits) {
    # round all numeric variables
    # x: data frame 
    # digits: number of digits to round
    numeric_columns <- sapply(x, mode) == 'numeric'
    x[numeric_columns] <-  round(x[numeric_columns], digits)
    x
}

habitatID <- round_df(habitatID, 2)
habitat<- round_df(habitat, 2)
S1S2Metrics<- round_df(S1S2Metrics, 4)


#write the summary files outs
write.table(habitatID, file= paste(outPath,'/',domain,site,'_geomorph_thalwegByHabitatID_',surveyDate,'.csv',sep = ''),sep = ',', col.names = TRUE, row.names = F, quote = T, na = '')

write.table(habitat, file= paste(outPath,'/',domain,site,'_geomorph_thalwegByHabitatType_',surveyDate,'.csv',sep = ''),sep = ',', col.names = TRUE, row.names = F, quote = T, na = '')

write.table(S1S2habitat, file= paste(outPath,'/',domain,site,'_geomorph_S1S2habitatIDs_',surveyDate,'.csv',sep = ''),sep = ',', col.names = TRUE, row.names = F, quote = T, na = '')

write.table(S1S2Metrics, file= paste(outPath,'/',domain,site,'_geomorph_S1S2Length_',surveyDate,'.csv',sep = ''),sep = ',', col.names = TRUE, row.names = F, quote = T, na = '')

write.table(longProfileDF, file= paste(outPath,'/',domain,site,'_geomorph_thalwegLongProfile_',surveyDate,'.csv',sep = ''),
sep = ',', col.names = TRUE, row.names = F, quote = T, na = '')

write.table(featSummary, file= paste(outPath,'/',domain,site,'_geomorph_featureCount_',surveyDate,'.csv',sep = ''), sep = ',', col.names = TRUE, row.names = F, quote = T, na = '')

write.table(surveyPtsDF, file= paste(outPath,'/',domain,site,'_geomorph_processedSurveyPts_',surveyDate,'.csv',sep = ''),sep = ',', col.names = TRUE, row.names = F, quote = T, na = '')
```

```{r create clean shapfiles for the kmz}
#use the original thalweg line that includes all sections of the braided channels
habitatOverlayKML <- raster::intersect(thalweg,rawHabitat)

#dissolve the line feature by habitat ID
thalwegKMZ <- raster::aggregate(habitatOverlayKML, c('habitatID'))

# grab the habitat information for the braided channel parts that were removed to calculate thalweg length
geoFeatKML<- geoFeaturesOnly[!(geoFeaturesOnly@data$temp %in% temp),]

#channels should be full loops again
plot(geoFeatKML)
plot(geoFeaturesRevised,add = TRUE)

#overlay the habitat information
geoFeatKMLHabitat<- raster::intersect(geoFeatKML, rawHabitat)

#add length (in m) 
geoFeatKMLHabitat@data$LengthHab <- gLength(geoFeatKMLHabitat, byid=TRUE)

#transform to a dataframe, add the braided channel length onto the appropriate habitatID
geoFeatKMLHabitat<- data.frame(geoFeatKMLHabitat, stringsAsFactors = FALSE)
geoFeatKMLHabitat$habitatID <- as.character(geoFeatKMLHabitat$habitatID)

#remove the total line from the habitat id dataframe
habitatID <- head(habitatID, -1)

for (a in 1:nrow(habitatID)){
  for (z in 1:nrow(geoFeatKMLHabitat)){
  if(habitatID$habitatID[a] == geoFeatKMLHabitat$habitatID[z]){
    habitatID$lengthM[a] <- geoFeatKMLHabitat$LengthHab[z] + habitatID$lengthM[a]
    }
  }
} 

#join thalwegKMZ shapefile to the habitatID dataframe to update the attribute fields
thalwegKMZ2 <- merge(thalwegKMZ,habitatID, by ='habitatID')

#get rid of the columns you don't need 
# all that should be left is habitat ID and length
thalwegKMZ2@data
thalwegKMZ2<- thalwegKMZ2[,-(3:8)]
thalwegKMZ2@data

#clean up the habitat polygon
#habitat, area_m2
habitatKMZ <- merge(rawHabitat,habitatID, by ='habitatID')
plot(habitatKMZ)
habitatKMZ@data
habitatKMZ<- habitatKMZ[,-c(2:3,5:6)]
habitatKMZ@data

#clean up the S1S2 points shapefile, add the staff gauge
#name, latitude, longitude, northing, easting, elevation
S1S2 <- surveyPts[grepl("^S1$|^S2$",surveyPts@data$name),]
staffGauge <- surveyPts[grepl("SP_",surveyPts@data$name),] # no S1/S2 at REDB  sensorsKMZ <- staffGauge
sensorsKMZ <- rbind(S1S2,staffGauge)
plot(sensorsKMZ)
sensorsKMZ@data

#clean up geoMorphicFeatures (polygons)
#feature, area_m2
if (file.exists(paste(paste(filePath,geoFeaturesShapefileName, sep="/"),".shp",sep=""))){
  plot(geoFeatures)
  geoFeatures@data
  geoFeaturesKMZ<- geoFeatures[,-(3)]
  geoFeaturesKMZ@data
}

#clean up geoMorphicFeatures (points)
#name, latitude, longitude, northing, easting, elevation
geoFeatPointsKMZ <- surveyPts[grepl("WF|CUL|TRB|MAR",surveyPts@data$name),]
if (nrow(geoFeatPointsKMZ) > 0){
  geoFeatPointsKMZ@data
}

#pull out transects
#Name, northing, easting, lat, long, elevation, mapCode
transectsKMZ <- surveyPts[grepl("Transect",surveyPts@data$mapCode),] 
plot(transectsKMZ)
transectsKMZ@data

#export shapefiles
#this code does not overwrite existing shapefiles and so you will need to delete old files first
writeOGR(thalwegKMZ2, outPath, paste(domain,site,'_geomorph_thalweg_',surveyDate, sep = ''), "ESRI Shapefile")
writeOGR(habitatKMZ, outPath, paste(domain,site,'_geomorph_habitatUnits_',surveyDate, sep = ''), "ESRI Shapefile")
writeOGR(sensorsKMZ, outPath, paste(domain,site,'_geomorph_sensors_',surveyDate, sep = ''), "ESRI Shapefile")
writeOGR(transectsKMZ, outPath, paste(domain,site,'_geomorph_transects_',surveyDate, sep = ''), "ESRI Shapefile")

if (file.exists(paste(paste(filePath,geoFeaturesShapefileName, sep="/"),".shp",sep=""))){
  writeOGR(geoFeaturesKMZ, outPath, paste(domain,site,'_geomorph_geomorphicFeaturePolygons_',surveyDate, sep = ''), "ESRI Shapefile")
}

if (nrow(geoFeatPointsKMZ) > 0){
  writeOGR(geoFeatPointsKMZ, outPath, paste(domain,site,'_geomorph_geomorphicFeaturePoints_',surveyDate, sep = ''), "ESRI Shapefile")
}
```
