---
output:
  html_document: default
  pdf_document: default
---
##############################################################################################
@title 
geomorphTransectMetrics

@author
Nick Harrison, Research Scientist, National Ecological Observatory Network 
\email{nharrison@BattelleEcology.org}

@description 
This script provides the following metrics based on data collected during the NEON geomorphology protocol:

1. Bankfull widths for each stream transect where both right and left bankfull indicators were mapped.
2. Average bankfull width of all transects mapped.
3. Plots of each transects mapped.  Plot images are not read out as outputs.  

@parmam
The only input is a single file, DOMSITE_geomorph_processedSurveyPts_YYYMMDD.csv, derived using geomorphThalwegMetrics.RMD.  This file contains geo-referenced total station survey data.  

@return
1.One .CSV file that provides bankfull widths for each transect (where left and right bankfull indicators were mapped) and the average bankfull width of all transects.  If bankfull measurements (mapCodes ="RBF" or "LBF") are not available for a given transect, an "NA" value is returned.  

@changelog
N. Harrison (2017-10-18): original creation, .R format 
N. Harrison (2017-12-18): conversion to .RMD format
##############################################################################################

```{r Required packages}
#load packages
require(calibrate)
```

```{r Set variables}
domainID <- 'D07' #NEON Domain number (ex: D01)
siteID <- 'WALK' #Four-digit NEON site code (ex: HOPB)
surveyDate<-'20171208' #The end date of the geomorphology survey (YYYYMMDD)
processedFile<-'D07WALK_GEOMORPH_processedSurveyPts_20171208.csv' #name of the .CSV that contains the post-processed survey data 
wdir<-paste('C:/Users/nharrison/Documents/GitHub/landWaterSoilIPT/streamMorpho/ScienceProcessingCode/R_Metrics',siteID,sep="/") 
#working directory where the processedFile is located on your network.  
```

```{r Reads in the post-processed geomorphology survey data}
allData<-read.csv(file=paste(wdir,processedFile,sep="/"),header=TRUE,stringsAsFactors = FALSE)
```

```{r Evaluates which transects were mapped during the survey}
#TRUE = the transect was mapped during the survey; FALSE = the transect was not mapped during the survey.

if("Transect_DSC1" %in% allData$mapCode){
  SurveyedDSC1<-TRUE
}else {
  SurveyedDSC1<-FALSE
}

if("Transect_DSC2" %in% allData$mapCode){
  SurveyedDSC2<-TRUE
}else {
  SurveyedDSC2<-FALSE
}

if("Transect_AP1" %in% allData$mapCode){
  SurveyedAP1<-TRUE
}else {
  SurveyedAP1<-FALSE
}

if("Transect_AP2" %in% allData$mapCode){
  SurveyedAP2<-TRUE
}else {
  SurveyedAP2<-FALSE
}

if("Transect_AP3" %in% allData$mapCode){
  SurveyedAP3<-TRUE
}else {
  SurveyedAP3<-FALSE
}

if("Transect_AP4" %in% allData$mapCode){
  SurveyedAP4<-TRUE
}else {
  SurveyedAP4<-FALSE
}

if("Transect_AP5" %in% allData$mapCode){
  SurveyedAP5<-TRUE
}else {
  SurveyedAP5<-FALSE
}

if("Transect_AP6" %in% allData$mapCode){
  SurveyedAP6<-TRUE
}else {
  SurveyedAP6<-FALSE
}

if("Transect_AP7" %in% allData$mapCode){
  SurveyedAP7<-TRUE
}else {
  SurveyedAP7<-FALSE
}

if("Transect_AP8" %in% allData$mapCode){
  SurveyedAP8<-TRUE
}else {
  SurveyedAP8<-FALSE
}

if("Transect_AP9" %in% allData$mapCode){
  SurveyedAP9<-TRUE
}else {
  SurveyedAP9<-FALSE
}

if("Transect_AP10" %in% allData$mapCode){
  SurveyedAP10<-TRUE
}else {
  SurveyedAP10<-FALSE
}

if("Transect_DSR" %in% allData$mapCode){
  SurveyedDSR<-TRUE
}else {
  SurveyedDSR<-FALSE
}

if("Transect_USR" %in% allData$mapCode){
  SurveyedUSR<-TRUE
}else {
  SurveyedUSR<-FALSE
}

if("Transect_S1" %in% allData$mapCode){
  SurveyedS1<-TRUE
}else {
  SurveyedS1<-FALSE
}

if("Transect_S2" %in% allData$mapCode){
  SurveyedS2<-TRUE
}else {
  SurveyedS2<-FALSE
}
```  

```{r Creates a dataframe for all discharge XS1 points sorted by northing values, if present.}
if(SurveyedDSC1==TRUE){
dischargePointsXS1=subset(allData,mapCode=="Transect_DSC1")
dischargePointsXS1<-dischargePointsXS1[
  with(dischargePointsXS1,order(dischargePointsXS1$northing)),
  ]
rownames(dischargePointsXS1)<-seq(length=nrow(dischargePointsXS1)) 
}
```

```{r Creates a data frame for all discharge XS2 points sorted by northing values, if present.}
if(SurveyedDSC2==TRUE){
  dischargePointsXS2=subset(allData,mapCode=="Transect_DSC2")
  dischargePointsXS2<-dischargePointsXS2[
    with(dischargePointsXS2,order(dischargePointsXS2$northing)),
    ]
  rownames(dischargePointsXS2)<-seq(length=nrow(dischargePointsXS2))
}
```

```{r Calculates bankfull width of discharge XS1 and XS2, if present.}
if(SurveyedDSC1==TRUE){
  DSCXS1Bankfull<-abs((dischargePointsXS1$northing[grepl("RBF",dischargePointsXS1$name)])-(dischargePointsXS1$northing[grepl("LBF",dischargePointsXS1$name)]))
} else {
  DSCXS1Bankfull<-"NA"
}

if(SurveyedDSC2==TRUE) {
  DSCXS2Bankfull<-abs((dischargePointsXS2$northing[grepl("RBF",dischargePointsXS2$name)])-(dischargePointsXS2$northing[grepl("LBF",dischargePointsXS2$name)]))
} else {
  DSCXS2Bankfull<-"NA"
}
```


```{r Creates a data frame for each additional transect and sorts data by northing values, if present.}
AP1Points=subset(allData,mapCode=="Transect_AP1") #Transect AP1 (aquatic plant transect 1)
AP1Points<-AP1Points[
  with(AP1Points,order(AP1Points$northing)),
  ]
rownames(AP1Points)<-seq(length=nrow(AP1Points))

AP2Points=subset(allData,mapCode=="Transect_AP2") #Transect AP2
AP2Points<-AP2Points[
  with(AP2Points,order(AP2Points$northing)),
  ]
rownames(AP2Points)<-seq(length=nrow(AP2Points)) 

AP3Points=subset(allData,mapCode=="Transect_AP3") #Transect AP3
AP3Points<-AP3Points[
  with(AP3Points,order(AP3Points$northing)),
  ]
rownames(AP3Points)<-seq(length=nrow(AP3Points))

AP4Points=subset(allData,mapCode=="Transect_AP4") #Transect AP4
AP4Points<-AP4Points[
  with(AP4Points,order(AP4Points$northing)),
  ]
rownames(AP4Points)<-seq(length=nrow(AP4Points))

AP5Points=subset(allData,mapCode=="Transect_AP5") #Transect AP5
AP5Points<-AP5Points[
  with(AP5Points,order(AP5Points$northing)),
  ]
rownames(AP5Points)<-seq(length=nrow(AP5Points))

AP6Points=subset(allData,mapCode=="Transect_AP6") #Transect AP6
AP6Points<-AP6Points[
  with(AP6Points,order(AP6Points$northing)),
  ]
rownames(AP6Points)<-seq(length=nrow(AP6Points))

AP7Points=subset(allData,mapCode=="Transect_AP7") #Transect AP7
AP7Points<-AP7Points[
  with(AP7Points,order(AP7Points$northing)),
  ]
rownames(AP7Points)<-seq(length=nrow(AP7Points))

AP8Points=subset(allData,mapCode=="Transect_AP8") #Transect AP8
AP8Points<-AP8Points[
  with(AP8Points,order(AP8Points$northing)),
  ]
rownames(AP8Points)<-seq(length=nrow(AP8Points))

AP9Points=subset(allData,mapCode=="Transect_AP9") #Transect AP9
AP9Points<-AP9Points[
  with(AP9Points,order(AP9Points$northing)),
  ]
rownames(AP9Points)<-seq(length=nrow(AP9Points))

AP10Points=subset(allData,mapCode=="Transect_AP10") #Transect AP10
AP10Points<-AP10Points[
  with(AP10Points,order(AP10Points$northing)),
  ]
rownames(AP10Points)<-seq(length=nrow(AP10Points))

DSRPoints=subset(allData,mapCode=="Transect_DSR") #Transect DSR (downstream reach boundary)
DSRPoints<-DSRPoints[
  with(DSRPoints,order(DSRPoints$northing)),
  ]
rownames(DSRPoints)<-seq(length=nrow(DSRPoints))

USRPoints=subset(allData,mapCode=="Transect_USR") #Transect USR (upstream reach boundary)
USRPoints<-USRPoints[
  with(USRPoints,order(USRPoints$northing)),
  ]
rownames(USRPoints)<-seq(length=nrow(USRPoints))

S1Points=subset(allData,mapCode=="Transect_S1") #Transect S1 (sensor set 1)
S1Points<-S1Points[
  with(S1Points,order(S1Points$northing)),
  ]
rownames(S1Points)<-seq(length=nrow(S1Points))

S2Points=subset(allData,mapCode=="Transect_S2") #Transect S2 (sensor set 2)
S2Points<-S2Points[
  with(S2Points,order(S2Points$northing)),
  ]
rownames(S2Points)<-seq(length=nrow(S2Points))
```

```{r Calculates bankfull widths for each transect based on right and left bankfull points.  Assigns an NA value to transects that were not surveyed.}
if(SurveyedAP1==TRUE){
  AP1Bankfull<-abs((AP1Points$northing[grepl("RBF",AP1Points$name)])-(AP1Points$northing[grepl("LBF",AP1Points$name)]))
} else {
  AP1Bankfull<-"NA"
}

if(SurveyedAP2==TRUE){
  AP2Bankfull<-abs((AP2Points$northing[grepl("RBF",AP2Points$name)])-(AP2Points$northing[grepl("LBF",AP2Points$name)]))
} else {
  AP2Bankfull<-"NA"
}

if(SurveyedAP3==TRUE){
  AP3Bankfull<-abs((AP3Points$northing[grepl("RBF",AP3Points$name)])-(AP3Points$northing[grepl("LBF",AP3Points$name)]))
} else {
  AP3Bankfull<-"NA"
}

if(SurveyedAP4==TRUE){
  AP4Bankfull<-abs((AP4Points$northing[grepl("RBF",AP4Points$name)])-(AP4Points$northing[grepl("LBF",AP4Points$name)]))
} else {
  AP4Bankfull<-"NA"
}

if(SurveyedAP5==TRUE){
  AP5Bankfull<-abs((AP5Points$northing[grepl("RBF",AP5Points$name)])-(AP5Points$northing[grepl("LBF",AP5Points$name)]))
} else {
  AP5Bankfull<-"NA"
}

if(SurveyedAP6==TRUE){
  AP6Bankfull<-abs((AP6Points$northing[grepl("RBF",AP6Points$name)])-(AP6Points$northing[grepl("LBF",AP6Points$name)]))
} else {
  AP6Bankfull<-"NA"
}

if(SurveyedAP7==TRUE){
AP7Bankfull<-abs((AP7Points$northing[grepl("RBF",AP7Points$name)])-(AP7Points$northing[grepl("LBF",AP7Points$name)]))
} else {
  AP7Bankfull<-"NA"
}

if(SurveyedAP8==TRUE){
AP8Bankfull<-abs((AP8Points$northing[grepl("RBF",AP8Points$name)])-(AP8Points$northing[grepl("LBF",AP8Points$name)]))
} else {
  AP8Bankfull<-"NA"
}

if(SurveyedAP9==TRUE){
AP9Bankfull<-abs((AP9Points$northing[grepl("RBF",AP9Points$name)])-(AP9Points$northing[grepl("LBF",AP9Points$name)]))
} else {
  AP9Bankfull<-"NA"
}

if(SurveyedAP10==TRUE){
AP10Bankfull<-abs((AP10Points$northing[grepl("RBF",AP10Points$name)])-(AP10Points$northing[grepl("LBF",AP10Points$name)]))
} else {
  AP10Bankfull<-"NA"
}

if(SurveyedS1==TRUE){
S1Bankfull<-abs((S1Points$northing[grepl("RBF",S1Points$name)])-(S1Points$northing[grepl("LBF",S1Points$name)]))
} else {
  S1Bankfull<-"NA"
}

if(SurveyedS2==TRUE){
S2Bankfull<-abs((S2Points$northing[grepl("RBF",S2Points$name)])-(S2Points$northing[grepl("LBF",S2Points$name)])) 
} else {
  S2Bankfull<-"NA"
}

if(SurveyedUSR==TRUE){
USRBankfull<-abs((USRPoints$northing[grepl("RBF",USRPoints$name)])-(USRPoints$northing[grepl("LBF",USRPoints$name)]))
} else {
  USRBankfull<-"NA"
}

if(SurveyedDSR==TRUE){
DSRBankfull<-abs((DSRPoints$northing[grepl("RBF",DSRPoints$name)])-(DSRPoints$northing[grepl("LBF",DSRPoints$name)]))
} else {
  DSRBankfull<-"NA"
}
```

```{r Assigns an NA bankfull width value to transects that were surveyed but did not contain both a right bankfull (RBF) and left bankfull (LBF) mapped point.}

if(length(AP1Bankfull)<1){
  AP1Bankfull<-"NA"
}

if(length(AP2Bankfull)<1){
  AP2Bankfull<-"NA"
}

if(length(AP3Bankfull)<1){
  AP3Bankfull<-"NA"
}

if(length(AP4Bankfull)<1){
  AP4Bankfull<-"NA"
}

if(length(AP5Bankfull)<1){
  AP5Bankfull<-"NA"
}

if(length(AP6Bankfull)<1){
  AP6Bankfull<-"NA"
}

if(length(AP7Bankfull)<1){
  AP7Bankfull<-"NA"
}

if(length(AP8Bankfull)<1){
  AP8Bankfull<-"NA"
}

if(length(AP9Bankfull)<1){
  AP9Bankfull<-"NA"
}

if(length(AP10Bankfull)<1){
  AP10Bankfull<-"NA"
}

if(length(S1Bankfull)<1){
  S1Bankfull<-"NA"
}

if(length(S2Bankfull)<1){
  S2Bankfull<-"NA"
}

if(length(USRBankfull)<1){
  USRBankfull<-"NA"
}

if(length(DSRBankfull)<1){
  DSRBankfull<-"NA"
}
```

```{r Creates a data frame that provides the bankfull width for each transect as well and the average bankfull width across all transects measured.}
transectBankfullWidths<-data.frame(matrix(nrow=1,ncol=17))
names(transectBankfullWidths)=c("AP1","AP2","AP3","AP4","AP5","AP6","AP7","AP8","AP9","AP10","USR","DSR","S1","S2","DSC_XS1","DSC_XS2","Average")
```

```{r Loads bankfull width data into data frame for each transect.  Averages bankfull width for all transects (excludes NA values).}
transectBankfullWidths$AP1<-as.numeric(AP1Bankfull)
transectBankfullWidths$AP2<-as.numeric(AP2Bankfull)
transectBankfullWidths$AP3<-as.numeric(AP3Bankfull)
transectBankfullWidths$AP4<-as.numeric(AP4Bankfull)
transectBankfullWidths$AP5<-as.numeric(AP5Bankfull)
transectBankfullWidths$AP6<-as.numeric(AP6Bankfull)
transectBankfullWidths$AP7<-as.numeric(AP7Bankfull)
transectBankfullWidths$AP8<-as.numeric(AP8Bankfull)
transectBankfullWidths$AP9<-as.numeric(AP9Bankfull)
transectBankfullWidths$AP10<-as.numeric(AP10Bankfull)
transectBankfullWidths$USR<-as.numeric(USRBankfull)
transectBankfullWidths$DSR<-as.numeric(DSRBankfull)
transectBankfullWidths$S1<-as.numeric(S1Bankfull)
transectBankfullWidths$S2<-as.numeric(S2Bankfull)
transectBankfullWidths$DSC_XS1<-as.numeric(DSCXS1Bankfull)
transectBankfullWidths$DSC_XS2<-as.numeric(DSCXS2Bankfull)
transectBankfullWidths$Average<-((rowSums(transectBankfullWidths,na.rm=TRUE)/length(transectBankfullWidths[1,][!is.na(transectBankfullWidths[1,])]))) 
```

```{r Creates an output data frame that lists bankfull widths for all transects.}
dataOut<-data.frame(matrix(nrow=17,ncol=2))
names(dataOut)=c("Transect","Bankfull Width (m)")
dataOut$Transect<-names(transectBankfullWidths)
dataOut[1,2]<-transectBankfullWidths$AP1
dataOut[2,2]<-transectBankfullWidths$AP2
dataOut[3,2]<-transectBankfullWidths$AP3
dataOut[4,2]<-transectBankfullWidths$AP4
dataOut[5,2]<-transectBankfullWidths$AP5
dataOut[6,2]<-transectBankfullWidths$AP6
dataOut[7,2]<-transectBankfullWidths$AP7
dataOut[8,2]<-transectBankfullWidths$AP8
dataOut[9,2]<-transectBankfullWidths$AP9
dataOut[10,2]<-transectBankfullWidths$AP10
dataOut[11,2]<-transectBankfullWidths$USR
dataOut[12,2]<-transectBankfullWidths$DSR
dataOut[13,2]<-transectBankfullWidths$S1
dataOut[14,2]<-transectBankfullWidths$S2
dataOut[15,2]<-transectBankfullWidths$DSC_XS1
dataOut[16,2]<-transectBankfullWidths$DSC_XS2
dataOut[17,2]<-transectBankfullWidths$Average
dataOut[,2]<-round(dataOut$`Bankfull Width (m)`,2)
```

```{r Writes out the output data frame into the working directory.}
write.csv(dataOut,paste(wdir,"/",domainID,siteID,"_geomorph_transectBankfullWidths_",surveyDate,".csv",sep=""),row.names=F)
```

```{r Plots available data for discharge XS1 and XS2 transects}

library(calibrate)

if(SurveyedDSC1==TRUE){
plot(dischargePointsXS1$northing,dischargePointsXS1$elevation,main=paste(siteID,"Discharge XS1 Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(dischargePointsXS1$northing,dischargePointsXS1$elevation,labels=dischargePointsXS1$name,pos=4)
lines(lines(dischargePointsXS1$northing,dischargePointsXS1$elevation,lty=3))
}

if(SurveyedDSC2==TRUE) {
plot(dischargePointsXS2$northing,dischargePointsXS2$elevation,main=paste(siteID,"Discharge XS2 Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(dischargePointsXS2$northing,dischargePointsXS2$elevation,labels=dischargePointsXS2$name,pos=4)
lines(lines(dischargePointsXS2$northing,dischargePointsXS2$elevation,lty=3))
}
```

```{r Plots available data for AP 1-4 transects (northing vs. elevation) in grouped windows.}
par(mfrow=c(2,2))

if(SurveyedAP1==TRUE){
plot(AP1Points$northing,AP1Points$elevation,main=paste(siteID,"AP1 XS Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(AP1Points$northing,AP1Points$elevation,labels=AP1Points$name,pos=4)
lines(lines(AP1Points$northing,AP1Points$elevation,lty=3))
}

if(SurveyedAP2==TRUE){
plot(AP2Points$northing,AP2Points$elevation,main=paste(siteID,"AP2 XS Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(AP2Points$northing,AP2Points$elevation,labels=AP2Points$name,pos=4)
lines(lines(AP2Points$northing,AP2Points$elevation,lty=3))
}

if(SurveyedAP3==TRUE){
plot(AP3Points$northing,AP3Points$elevation,main=paste(siteID,"AP3 XS Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(AP3Points$northing,AP3Points$elevation,labels=AP3Points$name,pos=4)
lines(lines(AP3Points$northing,AP3Points$elevation,lty=3))
}

if(SurveyedAP4==TRUE){
plot(AP4Points$northing,AP4Points$elevation,main=paste(siteID,"AP4 XS Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(AP4Points$northing,AP4Points$elevation,labels=AP4Points$name,pos=4)
lines(lines(AP4Points$northing,AP4Points$elevation,lty=3))
}
```

```{r Plots available data for AP 5-8 transects (northing vs. elevation) in grouped windows.}
par(mfrow=c(2,2))

if(SurveyedAP5==TRUE){
plot(AP5Points$northing,AP5Points$elevation,main=paste(siteID,"AP5 XS Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(AP5Points$northing,AP5Points$elevation,labels=AP5Points$name,pos=4)
lines(lines(AP5Points$northing,AP5Points$elevation,lty=3))
}

if(SurveyedAP6==TRUE){
plot(AP6Points$northing,AP6Points$elevation,main=paste(siteID,"AP6 XS Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(AP6Points$northing,AP6Points$elevation,labels=AP6Points$name,pos=4)
lines(lines(AP6Points$northing,AP6Points$elevation,lty=3))
}

if(SurveyedAP7==TRUE){
plot(AP7Points$northing,AP7Points$elevation,main=paste(siteID,"AP7 XS Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(AP7Points$northing,AP7Points$elevation,labels=AP7Points$name,pos=4)
lines(lines(AP7Points$northing,AP7Points$elevation,lty=3))
}

if(SurveyedAP8==TRUE){
plot(AP8Points$northing,AP8Points$elevation,main=paste(siteID,"AP8 XS Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(AP8Points$northing,AP8Points$elevation,labels=AP8Points$name,pos=4)
lines(lines(AP8Points$northing,AP8Points$elevation,lty=3))
}
```

```{r Plots available data for AP 9-10 transects (northing vs. elevation) in grouped windows.}
par(mfrow=c(1,2))

if(SurveyedAP9==TRUE){
plot(AP9Points$northing,AP9Points$elevation,main=paste(siteID,"AP9 XS Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(AP9Points$northing,AP9Points$elevation,labels=AP9Points$name,pos=4)
lines(lines(AP9Points$northing,AP9Points$elevation,lty=3))
}

if(SurveyedAP10==TRUE){
plot(AP10Points$northing,AP10Points$elevation,main=paste(siteID,"AP10 XS Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(AP10Points$northing,AP10Points$elevation,labels=AP10Points$name,pos=4)
lines(lines(AP10Points$northing,AP10Points$elevation,lty=3))
}
```


```{r Plots available data for S1, S2, USR, and DSR transects (northing vs. elevation) in grouped windows.}
par(mfrow=c(2,2))

if(SurveyedS1==TRUE){
plot(S1Points$northing,S1Points$elevation,main=paste(siteID,"S1 XS Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(S1Points$northing,S1Points$elevation,labels=S1Points$name,pos=4)
lines(lines(S1Points$northing,S1Points$elevation,lty=3))
}

if(SurveyedS2==TRUE){
plot(S2Points$northing,S2Points$elevation,main=paste(siteID,"S2 XS Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(S2Points$northing,S2Points$elevation,labels=S2Points$name,pos=4)
lines(lines(S2Points$northing,S2Points$elevation,lty=3))
}


if(SurveyedUSR==TRUE){
plot(USRPoints$northing,USRPoints$elevation,main=paste(siteID,"USR XS Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(USRPoints$northing,USRPoints$elevation,labels=USRPoints$name,pos=4)
lines(lines(USRPoints$northing,USRPoints$elevation,lty=3))
}

if(SurveyedDSR==TRUE){
plot(DSRPoints$northing,DSRPoints$elevation,main=paste(siteID,"DSR XS Survey Data"),xlab="northing (m)",ylab="elevation (m)")
text(DSRPoints$northing,DSRPoints$elevation,labels=DSRPoints$name,pos=4)
lines(lines(DSRPoints$northing,DSRPoints$elevation,lty=3))
}
```

